---
title: "Econometrics- Interim Assignment 2 Notebook"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
# Clear the environment
rm(list = ls())  # Remove all objects from the environment

# Clear the console
cat("\014")  # Clears the console screen
```

# Initial MLR

```{r}
#######MLR#######

# Install packages
PackageNames <- c("reticulate", "tidyverse", "stargazer", "magrittr", "car", "foreign", "janitor", "lmtest", "GGally", "reshape2", "viridis")
for(i in PackageNames){
  if(!require(i, character.only = T)){
    install.packages(i, dependencies = T)
    require(i, character.only = T)
  }
}

#Import Data and Cleansing
{
  options("scipen" = 10)
  options()$scipen
  dat<- read.csv("C:/Users/rites/Music/Eco_Data.csv", header = TRUE)
  finances <- remove_empty(dat, which = c("rows", "cols"), cutoff = 1, quiet = TRUE)
  str(finances)
  colnames(finances) <- c('FY','Rnd','GrossProfit','NetSales','TotalOperatingExpense','OperatingIncome')
  print(finances)
  
}

#Converting all the values to numeric
{
# Convert all character columns to numeric
finances[] <- lapply(finances, function(x) if(is.character(x)) as.numeric(gsub(",", "", x)) else x)

# Check structure after conversion
str(finances)
}

# Estimate three different models
model1<- lm(OperatingIncome ~ GrossProfit, data=finances)
model2<- lm(OperatingIncome ~ GrossProfit + NetSales, data=finances)
model3<- lm(OperatingIncome ~ GrossProfit + NetSales + TotalOperatingExpense + Rnd, data=finances)
summary(model1)
summary(model2)
summary(model3)

# Load package and display table of results
library(stargazer)
stargazer(list(model1,model2,model3),type="text",keep.stat=c("n","rsq"))

# Calculating the Akaike Information Criterion (AIC)
AIC(model1)
AIC(model2)
AIC(model3)  

# Run Regression
model <- lm(OperatingIncome ~ GrossProfit + NetSales + TotalOperatingExpense + Rnd, finances)
summary(model)

# Hypothesis testing method 1: compare t-statistic with t-critical value(s)
(coefficient <- coef(model)["NetSales"])
(se <- vcov(model) %>% diag %>% sqrt %>% .["NetSales"])  #Standard error of 'NetSales'
(tstat <- coefficient / se)  # t-statistic

# Degrees of freedom (n-k-1)
(df_r <- model$df.residual)

# t-critical value at 5% significance level
qt(p = 0.975, df = df_r, lower.tail = TRUE)

# Hypothesis testing method 2: compare p-value with significance level (5%)
p_value <- 2 * pt(q = abs(tstat), df = df_r, lower.tail = FALSE)

# Hypothesis testing method 3: calculate confidence intervals
CT5 = qt(p = 0.975, df = df_r)  # Critical value at 5%
lower_bound <- coefficient - CT5 * se  # Lower bound at 95% confidence level
upper_bound <- coefficient + CT5 * se  # Upper bound at 95% confidence level

# F-test for single coefficient significance ---------------------------
model_r1 <- lm(OperatingIncome ~ GrossProfit + TotalOperatingExpense + Rnd, finances)  # Restricted model
(ssr_r1 <- sum(resid(model_r1)^2))  # SSR for the restricted model
(df_r1 <- model_r1$df.residual)
model_ur <- model# Unrestricted model
print(model_ur)
print(model_r1)
(ssr_ur <- sum(resid(model_ur)^2))  # SSR for the unrestricted model
(df_ur <- model_ur$df.residual)  # Degrees of freedom for unrestricted model

cat("\nF-stat\n")
# Calculate F-stat using ssr_r and ssr_ur
F_stat <- ((ssr_r1 - ssr_ur) / 1) / (ssr_ur / df_ur)
print(F_stat)
# Calculate F-critical value
cat("\nF-critical\n")
F_critical <- qf(p = 0.95, df1 = 1, df2 = df_ur)
print(F_critical)
# p-value for F-test
F_pvalue <- pf(q = F_stat, df1 = 1, df2 = df_ur, lower.tail = FALSE)
cat("\np-value\n")
print(F_pvalue)
```

# Checking for multicollinearity (excluding the dependent variable)

```{r}
X <- finances[, c('GrossProfit','NetSales','TotalOperatingExpense','Rnd')]

# Add an intercept column to match the regression model
X <- cbind(Intercept = 1, X)

# Compute VIF
vif_values <- vif(model)  # Using 'car' package function
print(vif_values)
```

# Column Changes due to high VIF implying multicolinearity

```{r}
finances$ProfitperSales <- ifelse(finances$NetSales == 0, NA, finances$GrossProfit/finances$NetSales)
finances$OpsExperSales <- ifelse(finances$NetSales == 0, NA, finances$TotalOperatingExpense/finances$NetSales)
finances$RndperExp <- ifelse(finances$TotalOperatingExpense == 0, NA, finances$Rnd/finances$TotalOperatingExpense)

str(finances)
```

# Removing Redundancy

```{r}
library(dplyr)

finances <- finances %>%
  select(-TotalOperatingExpense, -NetSales, -GrossProfit, -Rnd)  # Remove specific columns

```

# Modelling it again

```{r}
#######MLR#######

#Import Data and Cleansing
{
  options("scipen" = 10)
  options()$scipen
  str(finances)
  print(finances)
}

# Run Regression
model <- lm(OperatingIncome ~ ProfitperSales + OpsExperSales + RndperExp, finances)
summary(model)

# Hypothesis testing method 1: compare t-statistic with t-critical value(s)
(coefficient <- coef(model)["RndperExp"])
(se <- vcov(model) %>% diag %>% sqrt %>% .["RndperExp"])  #Standard error of 'NetSales'
(tstat <- coefficient / se)  # t-statistic

# Degrees of freedom (n-k-1)
(df_r <- model$df.residual)

# t-critical value at 5% significance level
qt(p = 0.975, df = df_r, lower.tail = TRUE)

# Hypothesis testing method 2: compare p-value with significance level (5%)
p_value <- 2 * pt(q = abs(tstat), df = df_r, lower.tail = FALSE)

# Hypothesis testing method 3: calculate confidence intervals
CT5 = qt(p = 0.975, df = df_r)  # Critical value at 5%
lower_bound <- coefficient - CT5 * se  # Lower bound at 95% confidence level
upper_bound <- coefficient + CT5 * se  # Upper bound at 95% confidence level

# F-test for single coefficient significance ---------------------------
model_r1 <- lm(OperatingIncome ~ ProfitperSales + OpsExperSales, finances)  # Restricted model
(ssr_r1 <- sum(resid(model_r1)^2))  # SSR for the restricted model
(df_r1 <- model_r1$df.residual)
model_ur <- model# Unrestricted model
print(model_ur)
print(model_r1)
(ssr_ur <- sum(resid(model_ur)^2))  # SSR for the unrestricted model
(df_ur <- model_ur$df.residual)  # Degrees of freedom for unrestricted model

cat("\nF-stat\n")
# Calculate F-stat using ssr_r and ssr_ur
F_stat <- ((ssr_r1 - ssr_ur) / 1) / (ssr_ur / df_ur)
print(F_stat)
# Calculate F-critical value
cat("\nF-critical\n")
F_critical <- qf(p = 0.95, df1 = 1, df2 = df_ur)
print(F_critical)
# p-value for F-test
F_pvalue <- pf(q = F_stat, df1 = 1, df2 = df_ur, lower.tail = FALSE)
cat("\np-value\n")
print(F_pvalue)
```

# Multicolinearity second check

```{r}
str(finances)
X <- finances[, c('ProfitperSales', 'OpsExperSales', 'RndperExp')]

# Add an intercept column to match the regression model
X <- cbind(Intercept = 1, X)

# Compute VIF
vif_values <- vif(model)  # Using 'car' package function
print(vif_values)
```

# Removing High VIF columns to remove multicolinearity

```{r}
library(dplyr)

finances <- finances %>%
  select(-ProfitperSales, -OpsExperSales)  # Remove specific columns
str(finances)
```

# Autocorrelation Check

```{r}
# 1. Durbin-Watson Test for autocorrelation
dw_test <- dwtest(model)
print("Durbin-Watson Test Result:")
print(dw_test)

# 2. Breusch-Godfrey Test for autocorrelation (Higher order)
bg_test <- bgtest(model, order = 1)
print("Breusch-Godfrey Test Result:")
print(bg_test)


```

# Since DW value=1.2, suggesting positive autocorrelation in residuals, but BG p-value is > 0.05 thus no Serial Correlation. This is Time Series data so it is expected, we can ignore this.



```{r}
# Load package
library(ggplot2)

# Plot ACF of residuals
acf(model$residuals, main="Autocorrelation of Residuals")


# Save DataFrame to CSV
write.csv(finances, "finances_data.csv", row.names = FALSE)
```

# Now we test for heteroskedascity

```{r}
# Correlation matrix
cor_matrix <- cor(finances[, c('RndperExp', 'OperatingIncome')])
print(cor_matrix)

########### Heteroskedasticity Tests ###########

# 1. Breusch-Pagan LM Test
cat("\nBreusch-Pagan Test\n")
bp_test <- bptest(model)
print(bp_test)

# 2. Glesjer LM Test (Correcting the residuals part)
cat("\nGlesjer Test\n")
glesjer_model <- lm(abs(residuals(model)) ~ RndperExp, data=finances)
summary(glesjer_model)

# 3. Harvey-Godfrey LM Test (log of squared residuals)
cat("\nHarvey-Godfrey Test\n")
hg_model <- lm(log(residuals(model)^2) ~ RndperExp, data=finances)
summary(hg_model)


# 4. Goldfeld-Quandt Test
cat("\nGoldfeld-Quandt Test\n")
gq_test <- gqtest(model, order.by = ~ RndperExp, data = finances)
print(gq_test)

# 5. White's Test (testing for heteroskedasticity using squared residuals)
cat("\nWhite's Test\n")
white_test <- bptest(model, ~ RndperExp + I(RndperExp^2), data = finances)
print(white_test)

########### Visualizing Residuals ###########
par(mfrow = c(1, 2))  # Split plotting area
{plot(finances$RndperExp, residuals(model), main = "Residuals vs RndperExp", xlab = "RndperExp", ylab = "Residuals")
abline(h = 0, col = "red")}
hist(residuals(model), main = "Histogram of Residuals", xlab = "Residuals", col = "blue", breaks = 15)


#Remedial measure for heteroskedasticity 

##WLS
# Estimate weights based on residuals from initial model
weights <- 1 / residuals(model)^2
# Fit weighted least squares model
wls_model <- lm(OperatingIncome ~ RndperExp, data=finances, weights = weights)
summary(wls_model)

##GLS
# Install nlme package if not already installed
#install.packages("nlme")
#library(nlme)
# Fit GLS model
#gls_model <- gls(OperatingIncome ~ Cash + GainLoss + RevperOpEx + FinperRaw + WIPperRaw, data=finances, weights = #varIdent(form = ~1|WIPperRaw))
#summary(gls_model)

#compare results 
summary(model)  # OLS model summary
#summary(gls_model)  # GLS model summary
summary(wls_model)
#AIC(gls_model)
AIC(wls_model)
```

# Homoskedasticity was found thus we proceed with plot generation as our model is accurate enough.

```{r}
# Apply log transformation
finances$Log_OperatingIncome <- log(finances$OperatingIncome)

# Fit the regression with transformed dependent variable
log_model <- lm(Log_OperatingIncome ~ RndperExp, data=finances)

summary(log_model)

```

#Rechecking for Heteroskedaasticity

```{r}
bptest(log_model)
```

#Re-performing MLR

```{r}

# Hypothesis testing method 1: compare t-statistic with t-critical value(s)
(coefficient <- coef(log_model)["GainLoss"])
(se <- vcov(log_model) %>% diag %>% sqrt %>% .["GainLoss"])  #Standard error of 'GainLoss'
(tstat <- coefficient / se)  # t-statistic

# Degrees of freedom (n-k-1)
(df_r <- log_model$df.residual)

# t-critical value at 5% significance level
qt(p = 0.975, df = df_r, lower.tail = TRUE)

# Hypothesis testing method 2: compare p-value with significance level (5%)
p_value <- 2 * pt(q = abs(tstat), df = df_r, lower.tail = FALSE)

# Hypothesis testing method 3: calculate confidence intervals
CT5 = qt(p = 0.975, df = df_r)  # Critical value at 5%
lower_bound <- coefficient - CT5 * se  # Lower bound at 95% confidence level
upper_bound <- coefficient + CT5 * se  # Upper bound at 95% confidence level

# F-test for single coefficient significance ---------------------------
model_r1 <- lm(Log_OperatingIncome ~ Cash + RevperOpEx + FinperRaw + WIPperRaw, finances)  # Restricted model
(ssr_r1 <- sum(resid(model_r1)^2))  # SSR for the restricted model
(df_r1 <- model_r1$df.residual)
model_ur <- log_model# Unrestricted model
print(model_ur)
print(model_r1)
(ssr_ur <- sum(resid(model_ur)^2))  # SSR for the unrestricted model
(df_ur <- model_ur$df.residual)  # Degrees of freedom for unrestricted model

cat("\nF-stat\n")
# Calculate F-stat using ssr_r and ssr_ur
F_stat <- ((ssr_r1 - ssr_ur) / 1) / (ssr_ur / df_ur)
print(F_stat)
# Calculate F-critical value
cat("\nF-critical\n")
F_critical <- qf(p = 0.95, df1 = 1, df2 = df_ur)
print(F_critical)
# p-value for F-test
F_pvalue <- pf(q = F_stat, df1 = 1, df2 = df_ur, lower.tail = FALSE)
cat("\np-value\n")
print(F_pvalue)
```

#Redoing Autocorrelation checks

```{r}
print(dwtest(log_model))
print(bgtest(log_model))
```

# Since dwtest value has decreased below acceptable range we need to address it by adding Autoregressive Terms

```{r}
residuals <- resid(log_model)
lag_residuals <- c(NA, residuals[-length(residuals)])  # Lag the residuals
finances$lag_residuals <- lag_residuals

# Fit a new model with the lagged residuals
log_model_with_lag <- lm(Log_OperatingIncome ~ Cash + GainLoss + RevperOpEx + FinperRaw + WIPperRaw + lag_residuals, data = finances)
summary(log_model_with_lag)


print(dwtest(log_model_with_lag))
print(bgtest(log_model_with_lag))
```

# Residual Plots

```{r}
# Residuals vs Fitted plot
library(ggplot2)

# Extract fitted values
fitted_values <- fitted(log_model_with_lag)

# Plot residuals vs fitted values
plot<-ggplot(data = data.frame(fitted_values, residuals = resid(log_model_with_lag)), aes(x = fitted_values, y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Residuals vs Fitted Values", x = "Fitted Values", y = "Residuals") +
  theme_minimal()
ggsave("resvfit.png", plot = plot, dpi = 300, width = 8, height = 6)

# Q-Q plot for residuals
plot<-ggplot(data = data.frame(residuals = resid(log_model_with_lag)), aes(sample = residuals)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Q-Q Plot of Residuals") +
  theme_minimal()
ggsave("qq.png", plot = plot, dpi = 300, width = 8, height = 6)
```

# Pairplot

```{r}
# Plot pairplot using GGally's ggpairs function
library(GGally)

# Select the relevant columns for the pairplot
data_for_pairplot <- finances[, c("Cash", "GainLoss", "RevperOpEx", "FinperRaw", "WIPperRaw")]

# Create the pairplot
plot<-ggpairs(data_for_pairplot, 
        title = "Pairplot of Variables",
        lower = list(continuous = wrap("points", alpha = 0.3)), 
        upper = list(continuous = wrap("cor", size = 5))) + 
  theme(axis.text = element_blank(),   # Hide axis text (numbers)
        axis.ticks = element_blank())
ggsave("pairplot.png", plot = plot, dpi = 300, width = 8, height = 6)
```

# Heatmap

```{r}
# Compute the correlation matrix
cor_matrix <- cor(data_for_pairplot)

# Plot the correlation heatmap using the 'corrplot' package
library(corrplot)

# Plot the heatmap
corrplot(cor_matrix, method = "circle", type = "lower", 
         tl.cex = 0.8, tl.col = "black", 
         addCoef.col = "black", number.cex = 0.7, 
         main = "Correlation Heatmap")

library(ggplot2)
library(reshape2)
library(viridis)
df_filtered <- finances[, !(names(finances) %in% c("FY", "lag_residuals"))]  # Exclude 'FY' and 'lag_residuals'
data <- cor(df_filtered[sapply(df_filtered, is.numeric)])  # Correlation matrix
data1 <- melt(data)  # Melt the correlation matrix for ggplot

# Create the heatmap
plot<-ggplot(data1, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_viridis(discrete = FALSE) + # Use continuous color scale
  geom_text(aes(label = round(value, 2)), color = "black", size = 4) +
  labs(title = "Correlation Heatmap",
       x = "Variable 1",
       y = "Variable 2") +
  theme_minimal() +  # Add minimal theme for better aesthetics
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("correlation_heatmap.png", plot = plot, dpi = 300, width = 8, height = 6)
```

# Plots

```{r}
# Visualizing coefficients with their confidence intervals
coefs <- coef(summary(log_model_with_lag))
conf_int <- confint(log_model_with_lag)

# Create a plot for the coefficients and their CI
coef_data <- data.frame(
  Term = rownames(coefs),
  Estimate = coefs[, "Estimate"],
  CI_Lower = conf_int[, 1],
  CI_Upper = conf_int[, 2]
)

plot<-ggplot(coef_data, aes(x = Term, y = Estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), width = 0.2) +
  labs(title = "Coefficients with Confidence Intervals", x = "Terms", y = "Estimate") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("coefficients.png", plot = plot, dpi = 300, width = 8, height = 6)
```
